---
title: "Reproducible Research "
author: "Anisha Pal"
date: "29 March 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reproducible Research Assignment

1.Code for reading in the dataset and/or processing the data
2.Histogram of the total number of steps taken each day
3.Mean and median number of steps taken each day
4.Time series plot of the average number of steps taken
5.The 5-minute interval that, on average, contains the maximum number of steps
6.Code to describe and show a strategy for imputing missing data
7.Histogram of the total number of steps taken each day after missing values are imputed
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report


## Code for reading in the dataset and/or processing the data
```{r, echo = TRUE}
df <- read.csv("activity.csv")

df$date <- as.POSIXct(df$date, "%Y-%m-%d")
wday <- weekdays(df$date)
df <- cbind(df,wday)

summary(df)
```
## Histogram steps
```{r, echo = TRUE}
steps <- with(df, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(steps) <- c("date", "steps")
hist(steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "darkblue")
dev.copy (png,filename="plot1.png")
dev.off ()
```

## Mean and Median of steps taken
```{r, echo = TRUE}
mean(steps$steps)
median(steps$steps)
```

## average number of steps taken
```{r, echo = TRUE}
average <- aggregate(df$steps, by=list(df$interval), FUN=mean, na.rm=TRUE)
names(average) <- c("interval", "mean")
plot(average$interval, average$mean, type = "l", xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")
dev.copy (png,filename="plot2.png")
dev.off ()
```

## 5-minute interval that, on average, contains the maximum number of steps
```{r, echo = TRUE}
average[which.max(average$mean), ]$interval
```

## Code to describe and show a strategy for imputing missing data and Histogram of the total number of steps taken each day after missing values are imputed
```{r, echo = TRUE}
sum(is.na(df$steps))
imputed_steps <- average$mean[match(df$interval, average$interval)]
df_imputed <- transform(df, steps = ifelse(is.na(df$steps), yes = imputed_steps, no = df$steps))
total_steps_imputed <- aggregate(steps ~ date, df_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")
hist(total_steps_imputed$daily_steps, xlab = "Total steps per day", main = "Total number of steps taken each day")
dev.copy (png,filename="plot3.png")
dev.off ()
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r, echo = TRUE}
library(ggplot2)
df$date <- as.Date(strptime(df$date, format="%Y-%m-%d"))
df$datetype <- sapply(df$date, function(x) {
        if (weekdays(x) == "Sábado" | weekdays(x) =="Domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
df_by_date <- aggregate(steps~interval + datetype, df, mean, na.rm = TRUE)
plot<- ggplot(df_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
dev.copy (png,filename="plot4.png")
dev.off ()

```

